---
title: "What Will You Listen To?"
author: "Kai, Michael, and James"
date: '2017-12-12'
output: ioslides_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## Motivation
Recomendation is one of the hottest topics in data science and machine learning. It can be used for recomending:

- movies on netflix;

- products on amazon; and

- songs on online music (Spotify or WSDM)

We wanted to explor this further.

## Objectives:

Can we tell if a user will repeatedly listen to a specific song in the 1 month time frame after first hearing the song. 

This can be used to  play songs that the user is likely to enjoy, thus increasing their satisfaction with WSDM.

## Correlation
```{r error=FALSE, message=FALSE, warning=FALSE, echo=FALSE,results='hide',fig.keep='all'}
library(data.table)
library(tidyverse)
setwd("~/Data_607_Final_Project/")
members = fread('members.csv')
extra_song = fread('song_extra_info.csv')
songs = fread('songs.csv')
train = fread('train.csv')

 members$registration_init_time <- gsub("(\\d{4})(\\d{2})(\\d{2})$",
                                        "\\1-\\2-\\3", 
                                        members$registration_init_time) %>% 
  as.Date() %>% as.integer()

members$expiration_date <-  gsub("(\\d{4})(\\d{2})(\\d{2})$",
                                        "\\1-\\2-\\3", 
                                        members$expiration_date) %>% 
  as.Date() %>% as.integer()

empty_as_na <- function(x){
    if("factor" %in% class(x)) x <- as.character(x) ## since ifelse wont work with factors
    ifelse(as.character(x)!="", x, NA)
}

songs <- songs %>% mutate_all(funs(empty_as_na))
songs$composer <- ifelse(is.na(songs$composer), songs$artist_name, songs$composer) 
songs$lyricist <- ifelse(is.na(songs$lyricist), songs$composer, songs$lyricist)

# Train Join
train <- merge(train, members, by = 'msno')
train <- merge(train, songs, by = 'song_id')

train <- train %>% mutate_all(funs(empty_as_na)) 


train[is.na(data)] <- -1


# This converts all data to numeric:
# In effect, it makes everythign categorical and then assigns a number to them. 
train <- as.data.table(train)
for (f in names(train)){
  if( class(train[[f]]) == "character"){
    train[is.na(train[[f]]), eval(f) := ""]
    train[, eval(f) := as.integer(
      as.factor(train[[f]]))]
  } else train[is.na(train[[f]]), eval(f) := -1]
}




for (n in names(train)){
  xbar <- mean(train[[n]])
  xsd <- sd(train[[n]])
  train[[n]] <- (train[[n]] - xbar)/xsd
}

library(corrplot)
m <- cor(x = train)
corrplot(m, method = "circle")
```

## Correlation 2
No insight into trends initially. 
How can we make predictions?

## XGBoost 
**XGBoost**
 
* Efficent
* Accurate
* Tree Based
* CART (**C**lasification **A**nd **R**egression **T**ree)

## Results 
**Accuracy: val-auc:0.767018**

**Expected: 0.5**

## Challenges:

Data Manipulation:

* Converting factor level data (e.g., user ID or song name) into numeric values. 

* Scaling values with (Standardization)

$x^{\prime}=\frac{x - \bar{x}}{\sigma}$
